State-space analysis of a continuous gravitational wave source with a pulsar timing array: inclusion of the pulsar terms

### 2024-02-20 manuscript/manuscript_2/paper.tex
version submitted to MNRAS

There is appreciable natural dispersion in the one-dimensional posterior medians between realisations. We quantify the degree of dispersion using the coefficient of variation,
\begin{eqnarray}
	CV = \mu_{\rm med}/\sigma_{\rm med}
\end{eqnarray}
 where $\mu_{\rm med}$ is the mean value of the 10 posterior medians and $\sigma_{\rm med}$ is their standard deviation. The minimum and maximum $CV$ are 0.2\% and 52\% for $\Omega$ and $\Phi_0$ respectively. The remaining parameters typically satisfy $CV$ $\lesssim 10 \%$. Analogous results for synthetic data with higher strain ($h_0 = 1 \times 10^{-12}$) are presented in Appendix \ref{sec:app_high_SNR} (see Figure \ref{fig:corner_high_snr_appendix}) for completeness. The dispersion between noise realisations decreases as $h_0$ increases, and the one-dimensional posteriors for $\iota$ and $h_0$ are more symmetric about the injected value. \newline

\subsection{GW phase at each pulsar} \label{sec:chi_estim}
It is important to check how well the pulsar-term phase $\chi^{(n)}$ is estimated, since the reparameterization involving $\chi^{(n)}$ is a key feature of our approach, as explained in Section \ref{sec:pulsar_term2}. In short, it turns out that $\chi^{(n)}$ is estimated accurately for all $n$., except under special circumstances. Figure \ref{fig:corner_plot_3} displays the results as a corner plot for the representative subset $\chi^{(1)}, \dots, \chi^{(5)}$ over the ten noise realisations, in the same style as Figure \ref{fig:corner_plot_1}. We do not display the corner plot for $\chi^{(6)}$, $\dots$, $\chi^{(47)}$ because it is too big. With the exception of the results for $\chi^{(2)}$ (discussed below), all histograms and contours are consistent with a unimodal joint posterior, which peaks near the known, injected values. There is no evidence for correlations between parameter pairs. Most phases are recovered unambiguously across all noise realisations; for the 47 $\chi^{(n)}$ parameters (i.e. not just the five plotted in Figure \ref{fig:corner_plot_3}), the injected value is  contained within the 90\% credible interval of the one-dimensional marginalized posteriors in 383 out of the 470 possible combinations of the 47 parameters and ten realizations. The next paragraph explains why 383 out of 470 is fewer than 90\%. \newline

Sometimes, albeit rarely, $\chi^{(n)}$ is not estimated consistently across multiple realizations for some $n$, e.g $\chi^{(2)}$ (third row, third column of Figure \ref{fig:corner_plot_3}), corresponding to PSR J0340+4130. Out of all the pulsars in the array, this object matches most closely the direction to the synthetic GW source, with $\boldsymbol{n} \cdot \boldsymbol{q}^{(2)} = -0.96$ and $\cos \chi^{(2)} \approx 0.07$. The ability to accurately infer $\chi^{(2)}$ improves with the signal-to-noise ratio (SNR). For example, for $h_0 = 1 \times 10^{-12}$, $\chi^{(2)}$ is recovered unambiguously; see Figure \ref{fig:corner_high_snr_appendix_chi} in Appendix \ref{sec:app_high_SNR}. \newline

@@

\begin{figure*}
	\includegraphics[width=\textwidth, height =\textwidth ]{images/corner_lowSNR_n4000}
	\caption[Caption of LOF]{Posterior distribution of the GW source parameters $\boldsymbol{\theta}_{\rm gw}$ for the representative system in Table \ref{tab:parameters_and_priors}, for 10 realisations of the noise processes, with curves coloured uniquely per realisation. The horizontal and vertical orange lines indicate the true injected values. The contours in the two-dimensional histograms mark the (0.5, 1, 1.5, 2)-$\sigma$ levels after marginalizing over all but two parameters. The one-dimensional histograms correspond to the joint posterior distribution marginalized over all but one parameter. The supertitles of the one-dimensional histograms record the median and the 0.16 and 0.84 quantiles of the posterior median realisation (in this context the median realisation is defined as follows: a posterior is generated for all 10 realisations; the medians of the posteriors are ranked in ascending order; the median of the ranked list is defined to correspond to the median realisation). We plot the scaled variables $10^9 \Omega$ (units: ${\rm rad \, s^{-1}}$) and $10^{15} h_0$. The Kalman filter and nested sampler estimate accurately all seven parameters in ${\boldsymbol{\theta}}_{\rm gw}$. The horizontal axes span a subset of the prior domain for all seven parameters. The known, injected value lies within the 90\% credible interval for 60 out of the 70 combinations of seven parameters and ten noise realizations. There is appreciable dispersion among the peaks of the one-dimensional posteriors: the maximum coefficient of variation is $52 \%$ for $\Phi_0$, and the minimum is $0.2 \%$ for $\Omega$.}
	\label{fig:corner_plot_1}
\end{figure*}

@@

We analyse 200 injections constructed by fixing $h_0 = 5 \times 10^{-15}$ and $\iota =1.0$ rad and drawing the remaining five elements of $\boldsymbol{\theta}_{\rm gw}$ randomly from the prior distributions defined in Table \ref{tab:parameters_and_priors}. We fix $h_0$ and $\iota$ in order to maintain an approximately constant SNR across the 200 injections. For each injection we compute the posterior distribution of ${\boldsymbol{\theta}}_{\rm gw}$. The static pulsar parameters $\boldsymbol{\theta'}_{\rm psr}$ are specified in Section \ref{sec:rep_example1} and Table \ref{tab:parameters_and_priors}. \newline

It is prohibitive to display corner plots analogous to Figure \ref{fig:corner_plot_1} for 200 injections and seven elements of $\boldsymbol{\theta}_{\rm gw}$. Therefore, we summarise the results with the aid of a parameter-parameter (PP) plot \citep{doi:10.1198/106186006X136976}. A PP plot displays the fraction of injections included within a given credible interval of the estimated posterior, as a function of the credible interval itself. In the ideal case of perfect recovery, the PP plot should be a diagonal line of unit slope. \newline

The PP results are displayed in Figure \ref{fig:parameter_space}. The shaded grey contours enclose the $1\sigma$, $2\sigma$, and $3\sigma$ significance levels for 200 injections. The curves for the five parameters, colour-coded in the legend, are approximately linear. Some parameters are consistently well estimated across the parameter domain, e.g. the blue curve for $\Omega$ lies everywhere within the 2$\sigma$ shaded region. Conversely some parameters, e.g. $\Phi_0$, stray outside the $1\sigma$ shaded region and show evidence of being slightly over-constrained; there are more injections contained within lower-value credible intervals than would be expected statistically, and fewer injections contained within higher-value credible intervals. This occurs due to the low SNR of the injected GW signal and the varying sensitivity of the specific PTA configuration as a function of sky position. More quantitatively, for the best estimated parameter $\Omega$ the injection is contained within the 90\% credible interval in 91\% of cases. For the worst estimated parameter $\Phi_0$, the injection is contained within the 90\% credible interval in 81\% of cases; the next paragraph explains why 81\% is less than the theoretical ideal 90\%.  \newline

@@

For the high-SNR results in Figure \ref{fig:corner_plot_compare_high}, the one-dimensional posteriors for $\mathcal{M}_{\rm Earth}$ are evidently biased, as was observed by K24, as well as \cite{Zhupulsarterms} and \cite{Chen2022}. Particular biases are observed in $\psi, \iota $ and $\alpha$, with the blue posterior displaced with respect to the orange injection line. For example, for $\iota$, the injected value is contained within the 90\% credible interval, but the median of the posterior of the Earth-term model is shifted from the injected value by 0.35 radians. The inclusion of the pulsar terms corrects for this bias. The green one-dimensional posteriors exhibit no bias and are generally symmetric about the injected value. Moreover, for every pair of parameters in ${\boldsymbol{\theta}}_{\rm gw}$ the injected values are contained within the 2-sigma contours for $\mathcal{M}_{\rm psr \& Earth}$. This is not true for $\mathcal{M}_{\rm Earth}$ where the injected values fall outside the 2-sigma contour for 15 out of the 21 parameter pairs.\newline

We define a relative error to quantify the accuracy of the one-dimensional posteriors with respect to the injected value, viz.
\begin{eqnarray}
\Delta_{\rm M}(\theta) = \frac{ \left| \underset{\theta}{\text{argmax }} p_{\rm M}\left(\theta| \boldsymbol{Y}\right) - \theta_{\rm inj}\right|}{\theta_{\rm inj}} \ . \label{eq:mean_rel_error}
\end{eqnarray}
In Equation \eqref{eq:mean_rel_error}, $p_{\rm M}\left(\theta| \boldsymbol{Y}\right)$ is the one-dimensional posterior for parameter $\theta$ returned by nested sampling (c.f. Equation \eqref{eq:posterior_distrib}). The subscript $\rm M \in \{ \text{Earth, psr\&Earth} \}$ indicates whether the posterior is estimated using Equation \eqref{eq:g_func_trig_earth} or Equation \eqref{eq:g_func_trig} respectively. $\theta_{\rm inj}$ denotes the true injection value. We note that $\Delta_{\rm M}(\theta)$ quantifies the accuracy of the estimates returned by the two models, i.e. the closeness of the most probable estimate of $\theta$ relative to $\theta_{\rm inj}$. In contrast, $\Delta_{\rm M}(\theta)$ does not quantify the uncertainty in the estimates. The error $\Delta_{\rm M}(\theta)$ for each parameter in Figure \ref{fig:corner_plot_compare_high} is summarised in Table \ref{tab:posterior_errors}. We find that the estimates are more accurate using the pulsar terms, i.e. $\Delta_{\rm Earth}(\theta) > \Delta_{\rm psr \& Earth}(\theta)$ for $\theta$ in $\boldsymbol{\theta}_{\rm gw}$. In some cases the difference is modest;  $\Omega$ is recovered with high accuracy by both models, with $\Delta_{\rm Earth}(\Omega)-\Delta_{\rm psr \& Earth}(\Omega) = 1.5 \times 10^{-5}$. The improvement from including the pulsar terms is largest for $\iota$, with $\Delta_{\rm Earth}(\iota) - \Delta_{\rm psr \& Earth}(\iota)  = 0.14$. Whilst we present only a single noise realisation in this section, the improvements from including the pulsar terms are found to be comparable across different realisations. \newline

For the low-SNR results in Figure \ref{fig:corner_plot_compare_low}, there is less improvement in the parameter estimates. Qualitatively, the green and blue contours and histograms mostly overlap, although the green contours are centred slightly better on the injected values. The root-mean-square relative error $\Delta_{\rm M}(\theta)$ for the low SNR results is reported in the lower half of Table \ref{tab:posterior_errors}. We see that the error is larger for every element in $\boldsymbol{\theta}_{\rm gw}$ than in the high-SNR case, as expected. The inclusion of the pulsar terms improve the estimates for five out of the seven static parameters; we find $\Delta_{\rm Earth}(\theta) > \Delta_{\rm psr \& Earth}(\theta)$ for all parameters except $\psi$ and $h_0$. However it is hard to draw strong conclusions; the improvements from including the pulsar terms vary randomly across different noise realisations. Whilst including the pulsar terms improves the accuracy of the parameter estimates at low SNR modestly, we show in Section \ref{sec:psr_v_earth_bayes} that it increases the detection probability. \newline 
	\begin{figure*}
	\includegraphics[width=\textwidth, height =\textwidth ]{images/corner_highSNR_compare}
	\caption{Posterior distribution in the form of a standard corner plot of the GW source parameters $\boldsymbol{\theta}_{\rm gw}$ for the representative system described in Table \ref{tab:parameters_and_priors}, with $h_0 = 1 \times 10^{-12}$, for a single realisation of the system noise. The blue curves are the posteriors calculated using the Earth-term model, Equation \eqref{eq:g_func_trig_earth}. The green curves are calculated by including the Earth term and pulsar terms, Equation \eqref{eq:g_func_trig}. The vertical and horizontal orange lines indicate the true injected values. The contours in the two-dimensional histograms denote the (0.5, 1, 1.5, 2)-$\sigma$ levels. The supertitles of the one-dimensional histograms record the posterior medians and the 0.16 and 0.84 quantiles of the green curves, i.e. the pulsar-term model. We plot the scaled variables $10^9 \Omega$ (units: rad s$^{-1}$) and $10^{12} h_0$. Some parameters (e.g. $\psi, \iota$) exhibit a bias when using the Earth-term model which disappears when the pulsar terms are included.}
	\label{fig:corner_plot_compare_high}
\end{figure*}

@@

\begin{table}
	\centering
	%	\resizebox{\textwidth}{!}{%
		%	\renewcommand{\arraystretch}{1.0} % Default value: 1
		\begin{tabular}{lcll}
			\toprule
			&$\theta$ & $\Delta_{\rm Earth}(\theta)$ & $\Delta_{\rm psr \& Earth}(\theta)$  \\
			\hline
			\multirow{7}{2mm}{High SNR} & $\Omega$       & $1.9 \times 10^{-5}$& $3.4 \times 10^{-6}$\\
			& $\Phi_0$ &$1.6 \times 10^{-2}$ &$2.2 \times 10^{-3}$ \\
			& $\psi$ &$3.9 \times 10^{-2}$ &$2.0 \times 10^{-4}$  \\
			& $\iota$ & $2.0 \times 10^{-1}$ & $6.1 \times 10^{-2}$  \\
			& $\delta$ &$1.9 \times10^{-3}$ &$2.0 \times 10^{-4}$  \\
			&$\alpha$ &$4.0 \times 10^{-2}$ &$2.8 \times10^{-4}$  \\
			&$h_0$ & $9.3 \times 10^{-2}$ &$4.9 \times 10^{-2}$  \\
			\hline
			\multirow{7}{2mm}{Low SNR} & $\Omega$       &  $4.8 \times 10^{-3}$ &$1.5 \times 10^{-3}$ \\
			& $\Phi_0$ &$1.4 \times 10^{0}$ & $3.3 \times 10^{-1}$  \\
			& $\psi$ &$5.3 \times 10^{-3}$ &$1.1 \times 10^{-2}$  \\
			& $\iota$ & $3.3 \times 10^{-1}$ & $2.1 \times 10^{-1}$  \\
			& $\delta$ & $1.3 \times 10^{-1}$ & $4.2 \times 10^{-2}$  \\
			&$\alpha$ & $2.5 \times 10^{-2}$ & $3.3 \times 10^{-3}$ \\
			&$h_0$ & $1.1 \times 10^{-1}$ &$1.6 \times 10^{-1}$ \\
			\bottomrule
		\end{tabular}
		\caption{Relative error $\Delta_{\rm M}(\theta)$, Equation \eqref{eq:mean_rel_error}, in the one-dimensional posteriors calculated using the Earth-term model (M = Earth, Equation \eqref{eq:g_func_trig_earth}) and the pulsar-term model (M = $\rm psr \& Earth$, Equation \eqref{eq:g_func_trig}) for $\theta \in \boldsymbol{\theta}_{\rm gw}$. The injected values are as summarised in Table \ref{tab:parameters_and_priors}. The top and bottom halves of the table contain the high-SNR ($h_0 =1 \times 10^{-12}$) and low-SNR ($h_0 = 5 \times 10^{-15}$) cases respectively. The psr\&Earth model is more accurate than the Earth model for all $\theta \in \boldsymbol{\theta}_{\rm gw}$ at high SNR, and five out of seven $\theta \in \boldsymbol{\theta}_{\rm gw}$ at low SNR.}
		\label{tab:posterior_errors}
	\end{table}

@@

In this section we compute the minimum detectable strain for the representative source in Table \ref{tab:parameters_and_priors}, using $\mathcal{M}_{\rm Earth}$ and $\mathcal{M}_{\rm psr \& Earth}$. We frame the problem of claiming a GW detection in terms of a Bayesian model selection procedure, following the lead of other PTA analyses \citep[e.g.][]{2023ApJ...951L...8A,2023arXiv230616214A,2023ApJ...951L...6R,2023RAA....23g5024X}. We define $\mathcal{M}_{\rm null}$ as the null model that assumes no GW exists in the data. This is equivalent to setting $g^{(n)}(t)=1$ in Equation \eqref{eq:g_func_trig}. The evidence integral $\mathcal{Z}$ returned by nested sampling, Equation \eqref{eq:model_evidence}, is the probability of the data $\boldsymbol{Y}$ given a model $\mathcal{M}_{\rm M}$. The support in the data for the presence of a GW signal, described by model $\mathcal{M}_{\rm M}$, over the absence of a GW signal is quantified via the Bayes factor
		\begin{equation}
			\beta_{\rm M} = \frac{\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_{\rm M})}{\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_{\rm null})} \ . \label{eq:bayes}
		\end{equation}
In this paper we consider $\rm M \in \left \{ \rm Earth, psr \& Earth \right \}$. \newline

>>>>

There is appreciable natural dispersion in the one-dimensional posterior medians between realisations. We quantify the degree of dispersion using the coefficient of variation,
\begin{eqnarray}
	CV = \mu_{\rm med}/\sigma_{\rm med}
\end{eqnarray}
 where $\mu_{\rm med}$ is the mean of the 10 posterior medians and $\sigma_{\rm med}$ is their standard deviation. The minimum and maximum $CV$ are 0.2\% and 52\% for $\Omega$ and $\Phi_0$ respectively. The remaining parameters typically satisfy $CV$ $\lesssim 10 \%$. Analogous results for synthetic data with higher strain ($h_0 = 1 \times 10^{-12}$) are presented in Appendix \ref{sec:app_high_SNR} (see Figure \ref{fig:corner_high_snr_appendix}) for completeness. The dispersion between noise realisations decreases as $h_0$ increases, and the one-dimensional posteriors for $\iota$ and $h_0$ are more symmetric about the injected value. \newline

\subsection{GW phase at each pulsar} \label{sec:chi_estim}
It is important to check how well the pulsar-term phase $\chi^{(n)}$ is estimated, since the reparameterization involving $\chi^{(n)}$ is a key feature of our approach, as explained in Section \ref{sec:pulsar_term2}. In short, it turns out that $\chi^{(n)}$ is estimated accurately for all $n$, except under special circumstances. Figure \ref{fig:corner_plot_3} displays the results as a corner plot for the representative subset $\chi^{(1)}, \dots, \chi^{(5)}$ over the ten noise realisations, in the same style as Figure \ref{fig:corner_plot_1}. We do not display the corner plot for $\chi^{(6)}$, $\dots$, $\chi^{(47)}$ because it is too big. With the exception of the results for $\chi^{(2)}$ (discussed below), all histograms and contours are consistent with a unimodal joint posterior, which peaks near the known, injected values. There is no evidence for correlations between parameter pairs. Most phases are recovered unambiguously across all noise realisations; for the 47 $\chi^{(n)}$ parameters (i.e. not just the five plotted in Figure \ref{fig:corner_plot_3}), the injected value is  contained within the 90\% credible interval of the one-dimensional marginalized posteriors in 383 out of the 470 possible combinations of the 47 parameters and ten realizations. The next paragraph explains why 383 out of 470 is fewer than 90\%. \newline

Sometimes, albeit rarely, $\chi^{(n)}$ is not estimated consistently across multiple realizations for some $n$, e.g $\chi^{(2)}$ (third row, third column of Figure \ref{fig:corner_plot_3}) corresponding to PSR J0340+4130. Out of all the pulsars in the array, this object matches most closely the direction to the synthetic GW source, with $\boldsymbol{n} \cdot \boldsymbol{q}^{(2)} = -0.96$ and $\cos \chi^{(2)} \approx 0.07$. The ability to accurately infer $\chi^{(2)}$ improves with the signal-to-noise ratio (SNR). For example, for $h_0 = 1 \times 10^{-12}$, $\chi^{(2)}$ is recovered unambiguously; see Figure \ref{fig:corner_high_snr_appendix_chi} in Appendix \ref{sec:app_high_SNR}. \newline

@@

\begin{figure*}
	\includegraphics[width=\textwidth, height =\textwidth ]{images/corner_lowSNR_n4000}
	\caption[Caption of LOF]{Posterior distribution of the GW source parameters $\boldsymbol{\theta}_{\rm gw}$ for the representative system in Table \ref{tab:parameters_and_priors}, for 10 realisations of the noise processes, with curves coloured uniquely per realisation. The horizontal and vertical orange lines indicate the true injected values. The contours in the two-dimensional histograms mark the (0.5, 1, 1.5, 2)-$\sigma$ levels after marginalizing over all but two parameters. The one-dimensional histograms correspond to the joint posterior distribution marginalized over all but one parameter. The supertitles of the one-dimensional histograms record the median and the 0.16 and 0.84 quantiles of the median realisation. In this context the median realisation is defined as follows: a posterior is generated for all 10 realisations; the medians of the posteriors are ranked in ascending order; the median of the ranked list is associated with the median realisation. We plot the scaled variables $10^9 \Omega$ (units: ${\rm rad \, s^{-1}}$) and $10^{15} h_0$. The Kalman filter and nested sampler estimate accurately all seven parameters in ${\boldsymbol{\theta}}_{\rm gw}$. The horizontal axes span a subset of the prior domain for all seven parameters. The known, injected value lies within the 90\% credible interval for 60 out of the 70 combinations of seven parameters and ten noise realizations. There is appreciable dispersion among the peaks of the one-dimensional posteriors: the maximum coefficient of variation is $CV = 52 \%$ for $\Phi_0$, and the minimum is $CV = 0.2 \%$ for $\Omega$.}
	\label{fig:corner_plot_1}
\end{figure*}

@@

We analyse 200 injections constructed by fixing $h_0 = 5 \times 10^{-15}$ and $\iota =1.0$ rad and drawing the remaining five elements of $\boldsymbol{\theta}_{\rm gw}$ randomly from the prior distributions defined in Table \ref{tab:parameters_and_priors}. We fix $h_0$ and $\iota$ in order to maintain an approximately constant SNR across the 200 injections. For each injection we compute the posterior distribution of ${\boldsymbol{\theta}}_{\rm gw}$. The static pulsar parameters $\boldsymbol{\theta'}_{\rm psr}$ are specified in Section \ref{sec:rep_example1} and Table \ref{tab:parameters_and_priors}. It is prohibitive to display corner plots analogous to Figure \ref{fig:corner_plot_1} for 200 injections and seven elements of $\boldsymbol{\theta}_{\rm gw}$. Therefore, we summarise the results with the aid of a parameter-parameter (PP) plot \citep{doi:10.1198/106186006X136976}. A PP plot displays the fraction of injections included within a given credible interval of the estimated posterior, as a function of the credible interval itself. In the ideal case of perfect recovery, the PP plot should be a diagonal line of unit slope. \newline

The PP results are displayed in Figure \ref{fig:parameter_space}. The shaded grey contours enclose the $1\sigma$, $2\sigma$, and $3\sigma$ significance levels for 200 injections. The curves for the five parameters, colour-coded in the legend, are approximately linear. Some parameters are consistently well estimated across the parameter domain, e.g. the blue curve for $\Omega$ lies everywhere within the 2$\sigma$ shaded region. Conversely some parameters, e.g. $\Phi_0$, stray outside the $1\sigma$ shaded region and show evidence of being slightly over-constrained; there are more injections contained within lower-value credible intervals than would be expected statistically, and fewer injections contained within higher-value credible intervals. This occurs due to the low SNR of the injected GW signal and the varying sensitivity of the specific PTA configuration as a function of sky position. More quantitatively, for the best estimated parameter $\Omega$ the injection is contained within the 90\% credible interval in 91\% of cases. For the worst estimated parameter $\Phi_0$, the injection is contained within the 90\% credible interval in 81\% of cases. The next paragraph explains why 81\% is less than the theoretical ideal 90\%.  \newline

@@

For the high-SNR results in Figure \ref{fig:corner_plot_compare_high}, the one-dimensional posteriors for $\mathcal{M}_{\rm Earth}$ are biased, as was observed by K24, as well as \cite{Zhupulsarterms} and \cite{Chen2022}. Particular biases are observed in $\psi, \iota $ and $\alpha$, with the blue posterior displaced with respect to the orange injection line. For example, for $\iota$, the injected value is contained within the 90\% credible interval, but the median of the posterior of the Earth-term model is shifted from the injected value by 0.35 radians. The inclusion of the pulsar terms corrects for this bias. The green one-dimensional posteriors exhibit no bias and are generally symmetric about the injected value. Moreover, for every pair of parameters in ${\boldsymbol{\theta}}_{\rm gw}$, the injected values are contained within the 2-sigma contours for $\mathcal{M}_{\rm psr \& Earth}$. This is not true for $\mathcal{M}_{\rm Earth}$, where the injected values fall outside the 2-sigma contour for 15 out of the 21 parameter pairs.\newline

We define a relative error to quantify the accuracy of the one-dimensional posteriors with respect to the injected value. The relative error is defined to be the relative unsigned displacement of the mode of $p_{\rm M}\left(\theta| \boldsymbol{Y}\right)$ from the injection, viz.
\begin{eqnarray}
\Delta_{\rm M}(\theta) = \frac{ \left| \underset{\theta}{\text{argmax }} p_{\rm M}\left(\theta| \boldsymbol{Y}\right) - \theta_{\rm inj}\right|}{\theta_{\rm inj}} \ . \label{eq:mean_rel_error}
\end{eqnarray}
In Equation \eqref{eq:mean_rel_error}, $p_{\rm M}\left(\theta| \boldsymbol{Y}\right)$ is the one-dimensional posterior for parameter $\theta$ returned by nested sampling (c.f. Equation \eqref{eq:posterior_distrib}). The subscript $\rm M \in \{ \text{Earth, psr\&Earth} \}$ indicates whether the posterior is estimated using Equation \eqref{eq:g_func_trig_earth} or Equation \eqref{eq:g_func_trig} respectively. $\theta_{\rm inj}$ denotes the true injection value. We note that $\Delta_{\rm M}(\theta)$ quantifies the accuracy of the estimates returned by the two models, i.e. the closeness of the most probable estimate of $\theta$ relative to $\theta_{\rm inj}$. In contrast, $\Delta_{\rm M}(\theta)$ does not quantify the uncertainty in the estimates. The error $\Delta_{\rm M}(\theta)$ for each parameter in Figure \ref{fig:corner_plot_compare_high} is summarised in Table \ref{tab:posterior_errors}. We find that the estimates are more accurate using the pulsar terms, i.e. $\Delta_{\rm Earth}(\theta) > \Delta_{\rm psr \& Earth}(\theta)$ for $\theta \in \boldsymbol{\theta}_{\rm gw}$. In some cases the difference is modest;  $\Omega$ is recovered with high accuracy by both models, with $\Delta_{\rm Earth}(\Omega)-\Delta_{\rm psr \& Earth}(\Omega) = 1.5 \times 10^{-5}$. The improvement from including the pulsar terms is largest for $\iota$, with $\Delta_{\rm Earth}(\iota) - \Delta_{\rm psr \& Earth}(\iota)  = 0.14$. Whilst we present only a single noise realisation in this section, the improvements from including the pulsar terms are found to be comparable across different realisations. \newline

For the low-SNR results in Figure \ref{fig:corner_plot_compare_low}, there is less improvement in the parameter estimates. Qualitatively, the green and blue contours and histograms mostly overlap, although the green contours are centred slightly better on the injected values. The relative error $\Delta_{\rm M}(\theta)$ for the low-SNR results is reported in the lower half of Table \ref{tab:posterior_errors}. We see that the error is larger for every element in $\boldsymbol{\theta}_{\rm gw}$ than in the high-SNR case, as expected. The inclusion of the pulsar terms improve the estimates for five out of the seven static parameters; we find $\Delta_{\rm Earth}(\theta) > \Delta_{\rm psr \& Earth}(\theta)$ for all parameters except $\psi$ and $h_0$. However it is hard to draw strong conclusions; the improvements from including the pulsar terms vary randomly across different noise realisations. We show in Section \ref{sec:psr_v_earth_bayes} that including the pulsar terms increases the detection probability. \newline

\begin{figure*}
	\includegraphics[width=\textwidth, height =\textwidth ]{images/corner_highSNR_compare}
	\caption{Posterior distribution in the form of a standard corner plot of the GW source parameters $\boldsymbol{\theta}_{\rm gw}$ for the representative system described in Table \ref{tab:parameters_and_priors}, with $h_0 = 1 \times 10^{-12}$, for a single realisation of the system noise. The blue curves are the posteriors calculated using the Earth-term model, Equation \eqref{eq:g_func_trig_earth}. The green curves are calculated by including the Earth term and pulsar terms, Equation \eqref{eq:g_func_trig}. The vertical and horizontal orange lines indicate the true injected values. The contours in the two-dimensional histograms denote the (0.5, 1, 1.5, 2)-$\sigma$ levels. The supertitles of the one-dimensional histograms record the medians and the 0.16 and 0.84 quantiles of the green curves, i.e. the pulsar-term model. We plot the scaled variables $10^9 \Omega$ (units: rad s$^{-1}$) and $10^{12} h_0$. Some parameters (e.g. $\psi, \iota$) exhibit a bias when using the Earth-term model, which disappears when the pulsar terms are included.}
	\label{fig:corner_plot_compare_high}
\end{figure*}

@@

\begin{table}
	\centering
	%	\resizebox{\textwidth}{!}{%
		%	\renewcommand{\arraystretch}{1.0} % Default value: 1
		\begin{tabular}{lcll}
			\toprule
			&$\theta$ & $\Delta_{\rm Earth}(\theta)$ & $\Delta_{\rm psr \& Earth}(\theta)$  \\
			\hline
			\multirow{7}{2mm}{High SNR} & $\Omega$       & $1.9 \times 10^{-5}$& $3.4 \times 10^{-6}$\\
			& $\Phi_0$ &$1.6 \times 10^{-2}$ &$2.2 \times 10^{-3}$ \\
			& $\psi$ &$3.9 \times 10^{-2}$ &$2.0 \times 10^{-4}$  \\
			& $\iota$ & $2.0 \times 10^{-1}$ & $6.1 \times 10^{-2}$  \\
			& $\delta$ &$1.9 \times10^{-3}$ &$2.0 \times 10^{-4}$  \\
			&$\alpha$ &$4.0 \times 10^{-2}$ &$2.8 \times10^{-4}$  \\
			&$h_0$ & $9.3 \times 10^{-2}$ &$4.9 \times 10^{-2}$  \\
			\hline
			\multirow{7}{2mm}{Low SNR} & $\Omega$       &  $4.8 \times 10^{-3}$ &$1.5 \times 10^{-3}$ \\
			& $\Phi_0$ &$1.4 \times 10^{0}$ & $3.3 \times 10^{-1}$  \\
			& $\psi$ &$5.3 \times 10^{-3}$ &$1.1 \times 10^{-2}$  \\
			& $\iota$ & $3.3 \times 10^{-1}$ & $2.1 \times 10^{-1}$  \\
			& $\delta$ & $1.3 \times 10^{-1}$ & $4.2 \times 10^{-2}$  \\
			&$\alpha$ & $2.5 \times 10^{-2}$ & $3.3 \times 10^{-3}$ \\
			&$h_0$ & $1.1 \times 10^{-1}$ &$1.6 \times 10^{-1}$ \\
			\bottomrule
		\end{tabular}
		\caption{Relative error $\Delta_{\rm M}(\theta)$, Equation \eqref{eq:mean_rel_error}, in the mode of the one-dimensional posteriors calculated using the Earth-term model (M = Earth, Equation \eqref{eq:g_func_trig_earth}) and the pulsar-term model (M = $\rm psr \& Earth$, Equation \eqref{eq:g_func_trig}) for $\theta \in \boldsymbol{\theta}_{\rm gw}$. The injected values are summarised in Table \ref{tab:parameters_and_priors}. The top and bottom halves of the table contain the high-SNR ($h_0 =1 \times 10^{-12}$) and low-SNR ($h_0 = 5 \times 10^{-15}$) cases respectively. The psr\&Earth model is more accurate than the Earth model for all $\theta \in \boldsymbol{\theta}_{\rm gw}$ at high SNR, and five out of seven $\theta \in \boldsymbol{\theta}_{\rm gw}$ at low SNR.}
		\label{tab:posterior_errors}
	\end{table}

@@

In this section we compute the minimum detectable strain for the representative source in Table \ref{tab:parameters_and_priors}, using $\mathcal{M}_{\rm Earth}$ and $\mathcal{M}_{\rm psr \& Earth}$. We frame the detection problem in terms of a Bayesian model selection procedure, following the lead of other PTA analyses \citep[e.g.][]{2023ApJ...951L...8A,2023arXiv230616214A,2023ApJ...951L...6R,2023RAA....23g5024X}. We define $\mathcal{M}_{\rm null}$ as the null model that assumes no GW exists in the data. This is equivalent to setting $g^{(n)}(t)=1$ in Equation \eqref{eq:g_func_trig}. The evidence integral $\mathcal{Z}$ returned by nested sampling, Equation \eqref{eq:model_evidence}, is the probability of the data $\boldsymbol{Y}$ given a model $\mathcal{M}_{\rm M}$. The support in the data for the presence of a GW signal, described by model $\mathcal{M}_{\rm M}$, over the absence of a GW signal is quantified via the Bayes factor
		\begin{equation}
			\beta_{\rm M} = \frac{\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_{\rm M})}{\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_{\rm null})} \ . \label{eq:bayes}
		\end{equation}
In this paper we consider $\rm M \in \left \{ \rm Earth, psr \& Earth \right \}$. \newline
